---
title: "Final_Project_Code"
author: "Logan Rivera"
date: "2025-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```
```{r}
data = read_csv("Data/Model2020.csv")
```

```{r}
summary(data$age_median)
```



# Backwards elimination 
# Original model - Every independent variable 
```{r}
model1_every_independent = lm(reg_rate_all_pct ~
              Percent_Less_Than_9th +
              Percent_9_12_NoDiploma +
              Percent_HS_Diploma +
              Percent_College_NoDegree +
              Percent_Associates +
              Percent_Bachelors + 
              Percent_Graduate +
              Unemployment_Rate +
              Percent_Poverty +
              Median_Income +
              age_median +
              pct_under18 +
              pct_65plus +
              pct_male +
              pct_female +
              pct_hispanic 
              , data)
summary(model1_every_independent)
```


# Model that has solid p values and R squared, but bad colinearity
```{r}
model2_working_model = lm(reg_rate_all_pct ~
              Percent_Less_Than_9th +
  
              Percent_HS_Diploma +
              Percent_College_NoDegree +
              Percent_Associates +
              Percent_Bachelors + 
              Percent_Graduate +
              Unemployment_Rate +
              Percent_Poverty +
              
              pct_65plus +
              pct_male +
              pct_female +
              pct_hispanic 
      
              , data)
summary(model2_working_model)
```

```{r}
library(car)
```



# checking colinearity of model 2
```{r}
vif(model2_working_model)
```


# fixed multicolinearity problems (final model)
```{r}
model3_multicolinearity = lm(reg_rate_all_pct ~
              Percent_Less_Than_9th +
              Percent_9_12_NoDiploma +
              Percent_College_NoDegree +
              Percent_Associates +
              Percent_Bachelors + 
              Percent_Graduate +
              Unemployment_Rate +
              Percent_Poverty +
              pct_65plus +
              pct_male +
              pct_hispanic 
              , data)
summary(model3_multicolinearity)
```

# final colinearity values
```{r}
vif(model3_multicolinearity)
```


# checking model conditions 

#linearity (independent against residuals)
# independent #1 - Percent less than 9th 
```{r}
ggplot(model3_multicolinearity) + 
  geom_point(aes(x=Percent_Less_Than_9th, y=.resid))
```


# 2 - Percent_9_12_NoDiploma
```{r}
ggplot(model3_multicolinearity) + 
  geom_point(aes(x=Percent_9_12_NoDiploma, y=.resid))
```

# 3 - Percent_College_NoDegree
```{r}
ggplot(model3_multicolinearity) + 
  geom_point(aes(x=Percent_College_NoDegree, y=.resid))
```

# 4 - Percent_Associates        
```{r}
ggplot(model3_multicolinearity) + 
  geom_point(aes(x=Percent_Associates, y=.resid))
```

# 5 - Percent_Bachelors
```{r}
ggplot(model3_multicolinearity) + 
  geom_point(aes(x=Percent_Bachelors, y=.resid))
```

# 6 - Percent_Graduate
```{r}
ggplot(model3_multicolinearity) + 
  geom_point(aes(x=Percent_Bachelors, y=.resid))
```

# 7 - Unemployment_Rate
```{r}
ggplot(model3_multicolinearity) + 
  geom_point(aes(x=Unemployment_Rate, y=.resid))
```

# 8 - Percent_Poverty
```{r}
ggplot(model3_multicolinearity) + 
  geom_point(aes(x=Percent_Poverty, y=.resid))
```

# 9 - pct_65plus
```{r}
ggplot(model3_multicolinearity) + 
  geom_point(aes(x=pct_65plus, y=.resid))
```

# 10 - pct_male
```{r}
ggplot(model3_multicolinearity) + 
  geom_point(aes(x=pct_male, y=.resid))
```


# 11 - pct_hispanic
```{r}
ggplot(model3_multicolinearity) + 
  geom_point(aes(x=pct_hispanic, y=.resid))
```


# Independence (Order of data collection)

# ???


# Nearly Normal Residuals (distribution of residuals)
```{r}
ggplot(model3_multicolinearity)+
  geom_density(aes(x=.resid))
```


# Equal variability (Values of residuals against fitted values)

```{r}
ggplot(model3_multicolinearity) + 
  geom_point(aes(x=.fitted, y=.resid))

```

```{r}
library(skimr)
library(DescTools)

```


```{r}
variables <- c("Percent_Less_Than_9th", "Percent_9_12_NoDiploma", "Percent_College_NoDegree", "Percent_Associates", "Percent_Bachelors", "Percent_Graduate", "Unemployment_Rate", "Percent_Poverty", "pct_65plus", "pct_male", "pct_hispanic")

results <- list()

for (col in variables) {
  
  results[[col]] <- data %>%
    summarise(
      mean = mean(.data[[col]], na.rm = TRUE),
      median = median(.data[[col]], na.rm = TRUE),
      mode = Mode(.data[[col]]),
      range = max(.data[[col]], na.rm = TRUE) - min(.data[[col]], na.rm = TRUE),
      variance = var(.data[[col]], na.rm = TRUE),
      standard_dev = sd(.data[[col]], na.rm = TRUE)
    ) %>%
    pivot_longer(
      cols = everything(),
      names_to = "statistic",
      values_to = "value"
    ) %>%
    mutate(variable = col)
}

final_output <- bind_rows(results)
final_output
```

